Safe Memory Regions for Big Data Processing

### 2016-10-11 paper/type-inference.tex
type inference

\newcommand{\soln}{\eta}
\newcommand{\saturate}[1]{{#1}^*}
\newcommand{\myground}[1]{\saturate{#1}_g}
\newcommand{\consOf}[1]{\textsc{GenConstraint}(#1)}
\newcommand{\satC}{\saturate{C}}
\newcommand{\groundC}{\myground{C}}

@@

\paragraph{Constraint Solver.}
Our algorithm for solving a set of constraints $C$ consists of the following steps:
\begin{enumerate}
\item We use an iterative fixed point (saturation) algorithm to compute
$\saturate{C}$, the set of all constraints logically implied by $C$.
\item Let $\myground{C}$ denote the subset of all ground constraints
(\ie, constraints without any region variable or predicate variable)
in $\saturate{C}$. We check if $\myground{C}$ is valid in the theory of
partial-order. If $\myground{C}$ is valid, the solution $\soln$ is determined
as below. Otherwise, the constraint solver fails, as no valid solution exists.
\begin{itemize}
\item For any predicate variable $\varphi$, $\predSubstFn(\varphi)$ is defined
to be $\conj \{ \pi_1 \outlives \pi_2 \;|\; \pi_1, \pi_2 \in \regionConstants, \isvalid{\varphi}{\pi_1 \outlives \pi_2} \in \saturate{C} \}$.
\item For any region variable $\rho$, we define $\regionSubstFn(\rho)$ to be any element of the
set $\{ \pi \in \regionDeltaMap(\rho) \:|\; \isvalid{\ell}{\rho = \pi} \in \saturate{C} \}$ 
if this set is non-empty.
Otherwise, the solver fails.
%\item For any region variable $\rho$,  we are guaranteed that there exists a unique
% $\pi \in \regionConstants$  such that $\saturate{C}$ contains some constraint
%with a consequent $\rho = \pi$. We define $\regionSubstFn(\rho)$ to be $\pi$.
\end{itemize}
\end{enumerate}

\paragraph{Step 1.}
We compute $\saturate{C}$ by iteratively applying the following rules until a fixed
point is reached:

@@

Point~\ref{item:context} makes use of Lemmas~\ref{lemma:gc-is-decomposable} and~\ref{lemma:decomposition}.

\paragraph{Step 2.}
Checking the validity of a ground constraint is straightforward.
A ground constraint is of the form $\isvalid{\conj_{i \in I} \ell_i}{r}$, where each
$\ell_i$ and $r$ is an outlives-constraint.
Let $D$ denote $\{ \isvalid{}{\ell_i} \;|\; i \in I \}$.
The given constraint is valid iff $r$ belongs to $\saturate{D}$.

@@

% Correctness follows by establishing:
\begin{theorem}
  \label{thm:closure}
Let $C = \consOf{q}$.
Any assignment $\soln$ satisfies $C$ iff $\soln$ satisfies $\satC$.
%% (b) $\satC$ is satisfiable iff $\myground{C}$ is valid.
%% (c) For any region variable $\rho$, $\satC$ contains some constraint of the
%% form $\isvalid{\ell}{\rho = \pi}$.
\end{theorem}

Note: We likely need to add another rule to the saturation rules for computing
$\satC$, namely $\isvalid{\ell}{r} \in \satC \Rightarrow \isvalid{\ell \conj \phi}{r} \in \satC$.
We need this to propagate constraints from an outer scope into an inner scope.

\TODO{Need to establish that there are no independent variables with same name.}
\end{proof}

@@

\begin{theorem}
\label{thm:constraint-solver-sc}
Let $C = \consOf{q}$.
(Soundness) If the constraint solver succeeds, it will return an assignment $\soln$ 
that satisfies $C$.
(Completeness)
Conversely, if $C$ is satisfiable, then the constraint solver will succeed.
\end{theorem}

@@

\begin{itemize}
  \item explain: non-determinism in defining region-variable binding is immaterial.
  \item explain lemma 2's use case (a) and (b)
  \item notation: decomposable is dependent on C
  \item cleanup: syntax of constraints (region-constraint, validity constraint, etc.)
    and later usage (abduction constraint)
  \item adjust spacing in RULEs
  \item for function-calls: need basic typing from core language to introduce region parameters
  \item discuss computation of least-solution to constraints?
  \item use of a set to represent conjunctions
  \item equality is shorthand for two inequalities
  \item use of alpha-renaming, and its use to discharge not-in contraints;
  \item Note that we can also clarify that the well-formedness constraints are
    syntactic sugar from basic set constraints
  \item Prove: Progress: Every region-variable will be unified with some constant.
  \item Notation: should we use a different notation to distinguish a validity constraint
    from actual entailment?
  \item update abstract/intro to stress aspects of constraint solver
  \item Change static semantics to use method precondition, without class precondition
\end{itemize}

>>>>

\newcommand{\soln}{\eta}
\newcommand{\saturate}[1]{{#1}^*}
\newcommand{\myground}[1]{\saturate{#1}_g}
\newcommand{\consOf}[1]{\textsc{GenConstraint}(#1)}
\newcommand{\satC}{\saturate{C}}
\newcommand{\groundC}{\myground{C}}
\newcommand{\rhoC}{C_\rho}
\newcommand{\solveCon}[1]{\textsc{Solve}(#1)}

@@

\paragraph{Constraint Solver.}
The first step in our algorithm for solving a set of constraints $C$ uses an
iterative fixed point algorithm to compute a $\satC \supseteq C$ of
constraints by adding  a set of constraints that are logically implied by $C$.
% \paragraph{Step 1.}
We compute $\saturate{C}$ by iteratively applying the following rules until a fixed
point is reached:

@@

% Point~\ref{item:context} makes use of Lemmas~\ref{lemma:gc-is-decomposable} and~\ref{lemma:decomposition}.

\begin{theorem}
  \label{thm:closure}
Let $C = \consOf{q}$.
An assignment $\soln$ satisfies $C$ iff $\soln$ satisfies $\satC$.
\end{theorem}

One key point of the first step is that it has the effect of unifying region variables with the
region constants that they should be bound to in the solution, as explained below.

Note: We likely need to add another rule to the saturation rules for computing
$\satC$, namely $\isvalid{\ell}{r} \in \satC \Rightarrow \isvalid{\ell \conj \phi}{r} \in \satC$.
We need this to propagate constraints from an outer scope into an inner scope.
\end{proof}

The set $\satC$ makes it easy to identify a solution $\soln$ (if one exists), as below.
\begin{itemize}
\item For any predicate variable $\varphi$, $\predSubstFn(\varphi)$ is defined
to be $\conj \{ \pi_1 \outlives \pi_2 \;|\; \pi_1, \pi_2 \in \regionConstants, \isvalid{\varphi}{\pi_1 \outlives \pi_2} \in \saturate{C} \}$.
\item For any region variable $\rho$, we define $\regionSubstFn(\rho)$ to be any element of the
set $\{ \pi \in \regionConstants \:|\; \isvalid{\ell}{\rho = \pi} \in \saturate{C} \}$.
\end{itemize}

% \paragraph{Step 2.}

% \begin{enumerate}
% \item We use an iterative fixed point (saturation) algorithm to compute
% $\saturate{C}$, the set of all constraints logically implied by $C$.
% \item

The set $\satC$ also makes it easy to check if $C$ is satisfiable.
Let $\myground{C}$ denote the subset of all ground constraints
(\ie, constraints without any region variable or predicate variable)
in $\saturate{C}$. Let $\rhoC$ denote the subset of all wellformedness
constraints for region variables in $\satC$.
Define $\solveCon(C)$ as below.
% It is possible to show that $\soln$ satisfies $\satC4
% iff  $\myground{C}$ is valid in the theory of partial-order.

% \end{enumerate}

\[
\solveCon{C} = 
\begin{cases}
\textsc{Some} (\soln) & \textrm{if $\groundC$ is valid and $\soln$ satisfies $\rhoC$} \\
\textsc{None} & \textrm{otherwise}
\end{cases}
\]

\begin{theorem}
\label{thm:constraint-solver-sc}
Let $C = \consOf{q}$.
(Soundness) If $\solveCon{C} = \textsc{Some}(\soln)$, then $\soln$ satisfies $C$.
(Completeness) If $\solveCon{C} = \textsc{None}$, then $C$ is unsatisfiable.
\end{theorem}

Checking the validity of a ground constraint is straightforward.
A ground constraint is of the form $\isvalid{\conj_{i \in I} \ell_i}{r}$, where each
$\ell_i$ and $r$ is an outlives-constraint.
Let $D$ denote $\{ \isvalid{}{\ell_i} \;|\; i \in I \}$.
The given constraint is valid iff $r$ belongs to $\saturate{D}$.

@@

% Correctness follows by establishing:

@@

@@

\begin{itemize}
  \item explain: non-determinism in defining region-variable binding is immaterial.
  \item explain lemma 2's use case (a) and (b)
  \item notation: decomposable is dependent on C
  \item cleanup: syntax of constraints (region-constraint, validity constraint, etc.)
    and later usage (abduction constraint)
  \item adjust spacing in RULEs
  \item for function-calls: need basic typing from core language to introduce region parameters
  \item discuss computation of least-solution to constraints?
  \item use of a set to represent conjunctions
  \item equality is shorthand for two inequalities
  \item use of alpha-renaming, and its use to discharge not-in contraints;
\item also remind reader that variable names are unique
  \item Note that we can also clarify that the well-formedness constraints are
    syntactic sugar from basic set constraints
  \item Prove: Progress: Every region-variable will be unified with some constant.
  \item Notation: should we use a different notation to distinguish a validity constraint
    from actual entailment?
  \item update abstract/intro to stress aspects of constraint solver
  \item Change static semantics to use method precondition, without class precondition
\end{itemize}
